---
title: "Untitled"
author: "Patrick Ma"
date: "3/6/2020"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(readr)
library(tm)
library(ggplot2)
library(lubridate)
library(tidyverse)
library(tidytext)
library(dplyr)
library(purrr)
library(stringr)
library(igraph)
library(wordcloud2)
library(ggraph)
library(topicmodels)

hair_dryer <- read_tsv("~/Downloads/Problem_C_Data/hair_dryer.tsv")
microwave <- read_tsv("~/Downloads/Problem_C_Data/microwave.tsv")
pacifier <- read_tsv("~/Downloads/Problem_C_Data/pacifier.tsv")
#hair_dryer_ratio <- read_tsv("~/Desktop/2020 Spring/2020 C /hair_dryer_ratio.tsv")
```


```{r}
mean(hair_dryer$star_rating)
sd(hair_dryer$star_rating)
hist(hair_dryer$star_rating)
```


```{r}
mean(hair_dryer$helpful_votes)
median(hair_dryer$helpful_votes)
sd(hair_dryer$helpful_votes)

mean(hair_dryer$total_votes)
median(hair_dryer$total_votes)
sd(hair_dryer$total_votes)

hair_dryer_ratio = read_tsv("Problem_C_Data/hair_dryer_ratio.tsv")

# mean for helpful ratio when there is a comment
mean(na.omit(hair_dryer_ratio$helpful_total_ratio))
```

```{r}
# extract product that receives more than 50 reviews
# b/c the median after drop products only receive 1 review  is 44
tt <- table(hair_dryer_ratio$product_parent)
mean(tt)
median(subset(tt, tt>1))
hist(subset(tt, tt>1), breaks = 100)
hair_dryer_ratio <- subset(hair_dryer_ratio, product_parent %in% names(tt[tt > 50]))
length(hair_dryer_ratio$product_parent)
write_tsv(hair_dryer_ratio,"Problem_C_Data/hair_dryer_ratio.tsv")

```




```{r}
# x: rating , y: vote helpful ratio
head(hair_dryer_ratio)
not_na_idx = is.na(hair_dryer_ratio$helpful_total_ratio)==FALSE

hair_dryer_srtr<-aggregate(star_rating~helpful_total_ratio, hair_dryer_ratio[not_na_idx,] , mean)
ggplot(data = hair_dryer_srtr, aes(helpful_total_ratio,star_rating))+geom_point() + labs(title = "Scatterplot of mean star rating vs.  helpful ratio (excluding NA) for hair_dryer", y="mean star rating", x = "helpful ratio")
```

```{r}
hair_dryer_srtr<-aggregate(helpful_total_ratio~ star_rating, hair_dryer_ratio[not_na_idx,] , mean)
ggplot(data = hair_dryer_srtr, aes(star_rating,helpful_total_ratio))+geom_line() + labs(title = "mean helpful ratio (excluding NA) vs. star rating for hair_dryer", y="mean helpful ratio", x = "star rating")
```

* helpful ratio for 5-star rating is highest, while helpful ratio for 1-star rating is lowest

# Vine and verified

```{r}
sum(hair_dryer_ratio$vine == "Y")/nrow(hair_dryer_ratio)
hair_dryer_ratio_vine = subset(hair_dryer_ratio, hair_dryer_ratio$vine=="Y")
hair_dryer_ratio_notvine = subset(hair_dryer_ratio, hair_dryer_ratio$vine=="N")
mean(hair_dryer_ratio_vine$star_rating)
mean(hair_dryer_ratio_notvine$star_rating)
t.test(hair_dryer_ratio_vine$star_rating, hair_dryer_ratio_notvine$star_rating)
t.test(hair_dryer_ratio_vine$helpful_votes, hair_dryer_ratio_notvine$helpful_votes)
```

* 0.01560593 of people are vine

* vine people tend to rate higher for hair_dryer

```{r}
sum(hair_dryer_ratio$verified_purchase == "Y")/nrow(hair_dryer_ratio)
hair_dryer_ratio_verified = subset(hair_dryer_ratio, hair_dryer_ratio$verified_purchase=="Y")
hair_dryer_ratio_notverified = subset(hair_dryer_ratio, hair_dryer_ratio$verified_purchase=="N")

mean(hair_dryer_ratio_verified$star_rating)
mean(hair_dryer_ratio_notverified$star_rating)
t.test(hair_dryer_ratio_verified$star_rating, hair_dryer_ratio_notverified$star_rating)
t.test(hair_dryer_ratio_verified$helpful_votes, hair_dryer_ratio_notverified$helpful_votes)

```

* 0.8553618 of reviews are verified

* validated people tend to rate higher for hair_dryer


# delete not verified accounts, as their helpful votes are significantly lower than verified account
```{r}
hair_dryer_ratio = subset(hair_dryer_ratio, hair_dryer_ratio$verified_purchase=='Y')
sum(hair_dryer_ratio$verified_purchase=='Y')
nrow(hair_dryer_ratio)
```

```{r}
library(itsmr)

hair_dryer_ratio$review_date=(mdy(hair_dryer_ratio$review_date))
hair_dryer_ratio$review_month = as.Date(cut(hair_dryer_ratio$review_date, breaks = "month"))
hair_dryer_ts = data.frame(hair_dryer_ratio[,8],hair_dryer_ratio[,17])
head(hair_dryer_ts)
hair_dryer_ts2<-aggregate(star_rating ~ review_month, hair_dryer_ts , mean)
ggplot( data = hair_dryer_ts2, aes(review_month, star_rating)) + geom_line() + labs(title = "Time series of mean star rating every month vs. review month for hair_dryer", y="mean star rating every month", x = "review month")
```


# clean the corpus
```{r}
# clean the review body
hair_dryer_ratio$review_body = stripWhitespace(removePunctuation(hair_dryer_ratio$review_body))
hair_dryer_ratio$review_body = tolower(hair_dryer_ratio$review_body)

# clean the review title
hair_dryer_ratio$review_headline = stripWhitespace(removePunctuation(hair_dryer_ratio$review_headline))
hair_dryer_ratio$review_headline = tolower(hair_dryer_ratio$review_headline)

write_tsv(hair_dryer_ratio,"Problem_C_Data/hair_dryer_ratio.tsv")
```




# C.2.a


## sentiment analysis & viz the freq table

```{r}
# load in the libraries we'll need
library(tidyverse)
library(tidytext)
library(glue)
library(stringr)


# get the sentiments of 1 star to 5 stars
source("GetSentiments.R")
GetSentiments(hair_dryer_ratio)
```


```{r warning=FALSE}
source("CloudFreq.R")

CloudFreq(hair_dryer_ratio$review_body)
```

* From the review body, we can see that blow, dry, cord, powerful, dry appears frequently


# bigram
```{r bigramming function, echo=F}
bigramming <- function(data){
  cbigram <- data %>% unnest_tokens(bigram, review_body, token = 'ngrams', n = 2)
  cbigram_sep <- cbigram %>% separate(bigram, c('first', 'second'), sep = ' ')
  cbigram2 <- cbigram_sep %>% filter(!first %in% stop_words$word, !second %in% stop_words$word, !str_detect(first,      '\\d'), !str_detect(second, '\\d')) %>% unite(bigram, c(first, second), sep = ' ') 
  return(cbigram2)
}
```

```{r}
hair_dryer_ratio$review_body <- removeWords(hair_dryer_ratio$review_body,c("dryer","dryers"))
top_bigrams <- bigramming(hair_dryer_ratio) %>% mutate(Rating = factor(star_rating, levels <- c(5:1))) %>% mutate(bigram = factor(bigram, levels = rev(unique(bigram)))) %>% group_by(star_rating) %>% count(bigram, sort=TRUE) %>% top_n(10, n) %>% ungroup() 

top_bigrams  %>% ggplot(aes(bigram, n, fill = star_rating)) + geom_col(show.legend = FALSE) + facet_wrap(~star_rating, ncol = 3, scales = 'free') + labs(x=NULL, y = 'frequency') + ggtitle('Most Common Bigrams for hair dryer (by star_rating)') + coord_flip()
```

##Network Visualization
First, let's do a network visualization of the 5-star reviews. To do so, we will use the `igraph` and `ggraph` libraries to show a network that highlights the shared words within the most common bigrams.


```{r igraph, message = F, warning = F, echo=F}
fivestar <- hair_dryer_ratio %>% filter(star_rating >= (4))
fivebi <- bigramming(fivestar) %>% count(bigram, sort = TRUE)
head(fivestar)
five_graph <- fivebi %>% separate(bigram, c('first','second'), sep = ' ') %>% filter(n > 25) %>% graph_from_data_frame()

set.seed(9190)
ggraph(five_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point(color = 'orangered1', size = 3) +
  geom_node_text(aes(label = name), vjust = 1.1, hjust = 1.1) + scale_x_continuous(limits = c(2, 18)) +
  labs(title = "Network analysis of >=4 star review body in hair dryer")
```
```{r}
set.seed(9190)

onestar <- hair_dryer_ratio %>% filter(star_rating <= 2)
onebi <- bigramming(onestar) %>% count(bigram, sort = TRUE)
head(onestar)
one_graph <- onebi %>% separate(bigram, c('first','second'), sep = ' ') %>% filter(n > 5) %>% graph_from_data_frame()

ggraph(one_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point(color = 'blue', size = 3) +
   geom_node_text(aes(label = name), vjust = 1.1, hjust = 1.1) + scale_x_continuous(limits = c(2, 18)) +
  labs(title = "Network Analysis of Hairdryer Reviews with Rating <= 2 Stars")
```
```{r}
library(ngram)
#hair_dryer$wordc=rep(NA,8981)
for(i in 1:7506){
  hair_dryer_ratio$wordc[i]=wordcount(hair_dryer_ratio$review_body[i])
}
View(hair_dryer_ratio)
```

```{r}
hair_dryer_ratio$row_num <- seq.int(nrow(hair_dryer_ratio))
library(DescTools)
hair_dryer_ratio$row_num <- rev(hair_dryer_ratio$row_num)
transform(hair_dryer_ratio, hair_dryer_ratio$row_num)
```
```{r}
df2=data.frame(hair_dryer_ratio[,8])
library(itsmr)
View(df2)

df2$star_rating=as.double(df2$star_rating)
plotc(maunaloa_1009_0919)
write.csv(df2,file="/Users/hzn//Desktop/2020 SPRING/xx.csv")
df3<- read_csv("~/Desktop/2020 Spring/xx.csv")
df3=df3[,-1]
View(df3)
unlist(df3)
as.matrix(df3)
as.double(df3$star_rating)
df4=df3[1:100,]


acf(max1$star_rating,lag.max=40,type="correlation",plot=TRUE)
a=c("season",10)
resid1=Resid(max1$star_rating,a)
#plotc(resid1)
#acf(resid1,lag.max=40,type="correlation",plot=TRUE)

b=c("season",10, "trend",1)
resid2=Resid(max1$star_rating,b)

#plotc(resid2)
#acf(resid2,lag.max=40,type="correlation",plot=TRUE)

c=arma(resid2,3,3) 
resid4=Resid(resid2, a = c)
#acf(resid4,lag.max=40,type="correlation",plot=TRUE)
#test(resid4)
library(itsmr)
forecast(max1$star_rating,b,c,h=24,opt=3)

library(forecast)
library(stats)
max1=subset(hair_dryer_ratio,product_parent=="593915883")
for(i in 1:129){
  if(max1$star_rating[i]<=3){
    max1$star_rating[i]=0
  } else{
    max1$star_rating[i]=1
  }
}
plotc(max1$star_rating)
auto.arima(max1$star_rating, trace=TRUE)

plotc(max1$star_rating)
View(max1)
a=c("season",10)
resid1=Resid(max1$star_rating,a)
plotc(resid1)


predict(arima(max1$star_rating, order = c(3,0,3)), n.ahead = 30)
forecast(arima(max1$star_rating, order = c(3,0,3)),h=24,opt=2)
print(arima(max1$star_rating, order = c(3,0,3)))

# 把45星 123分开看 做logistic

plotc(max1$star_rating)

View(max1)

library(forecast)
futurVal=forecast(arima(max1$star_rating[300:500], order = c(3,0,3)),h=30, level=c(70))
plot(futurVal)

tsd=ts(max1$star_rating)
View(tsd)
```
```{r}
library(smooth)
max1=subset(hair_dryer_ratio,product_parent=="357308868")
a=es(max1$star_rating, h=18, holdout=TRUE, silent=FALSE)
print(a)

library(forecast)
fit_ses <- ses(max1$star_rating)
summary(fit_ses)
#Plot the forecasted values
plot(fit_ses)

fit_holt <- holt(max1$star_rating)
summary(fit_holt)
#Plot the forecasted values
plot(fit_holt)
print(fit_holt)
```

